---
title: "GO_Compare"
author: "Alex Gil"
date: "8/13/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Comparing GO terms from an old and a new analysis

file location: /gpfs/projects/RestGroup/mariana/GO/

## The old analysis was generated with JGI
  * GO ID:  
    - Aurli1_GeneCatalog_proteins_20120618_GO.tab

## The new analysis was generated with [sanspanz](http://ekhidna2.biocenter.helsinki.fi/sanspanz/)

  * GO ID:    
    - GO_predi_det_Aurli1_ekhidna2.txt

## Load files

```{r echo=T, messages='false', echo=TRUE}
library("tidyverse")
library("here")
library("knitr")

jgi <- read_delim(here("GO_comparison","Aurli1_GeneCatalog_proteins_20120618_GO.tab"),delim = "\t")
jgi_GO <- jgi %>% 
  select(goAcc) %>%
  unique() %>% 
  pull()

sans <- read_delim(here("GO_comparison","GO_predi_det_Aurli1_ekhidna2.txt"),delim = "\t")%>% rename(goAcc=goid) %>%
  mutate(goAcc=paste0("GO:",goAcc))
sans_GO <- sans  %>%
  select(goAcc) %>% 
  unique() %>% 
  pull()
```
# How many unique GO terms were found in each analysis?
```{r}
print("JGI")
jgi_GO  %>% length() #2156 unique Go_ids

print("Sanzpanz")
sans_GO  %>% length() #7234  unique Go_ids
```

# Which GO terms are in each set?
```{r}
tot <- c(sans_GO,jgi_GO) #8060 unique GO terms in total
first <- setdiff(jgi_GO,sans_GO) #826 unique GO terms in JGI
second <- setdiff(sans_GO,jgi_GO)  #5904 unique GO terms in Sans
both <- intersect(sans_GO,jgi_GO) #1330 unique GO terms in both
```

## Genes that were present in analysis 1, but not in 2.

```{r}
set1 <- jgi %>% filter(goAcc %in% first) %>% 
  select(goAcc,goName) %>%
  distinct() %>% 
  arrange(goAcc) 

set1
```


## Genes that were present in analysis 2, but not in 1.

```{r}
set2 <- sans %>% filter(goAcc %in% second) %>% 
  select(goAcc,desc) %>%
  distinct() %>% 
  arrange(goAcc)

set2
```

## Genes that were present in both analyses.

```{r}
set_both <- sans %>% filter(goAcc %in% both) %>% 
  select(goAcc,desc) %>%
  distinct() %>% 
  arrange(goAcc)

set_both
```

## Save results.
```{r}
write_csv(set1,here("GO_comparison","set1.csv"))
write_csv(set2,here("GO_comparison","set2.csv"))
write_csv(set_both,here("GO_comparison","set_both.csv"))
```

## Annotations

  * Is there a difference in the # of genes that got annotations?
  
  Yes, JGI got 6935 unique protein IDs, SansPanz got 5755 unique protein IDs.

```{r}
jgi_anno <- jgi %>% 
  rename(Protein_ID = `#proteinId`) %>% 
  select(Protein_ID) %>% 
  distinct() %>% 
  arrange(Protein_ID) %>% 
  pull() 
  

jgi_anno %>% length()  # 6935 protein IDs

sans_anno <- sans  %>%
  select(qpid) %>% 
  mutate(anno = str_split(qpid, "\\|",4)) %>% 
  mutate(Protein_ID = map(anno,nth,3)) %>% 
  mutate(Protein_ID=as.numeric(Protein_ID)) %>% 
  select(Protein_ID) %>% 
  distinct() %>% 
  arrange(Protein_ID) %>% 
  pull() 

sans_anno %>% length() # 5755 protein IDs
```

# Which annotations are in each set?
```{r}
tot <- c(sans_anno,jgi_anno) #8567 unique protein  IDs in total
first <- setdiff(jgi_anno,sans_anno) #2812 unique protein  IDs in JGI
second <- setdiff(sans_anno,jgi_anno)  #1632 unique protein  IDs in Sans
both <- intersect(sans_anno,jgi_anno) #4123 unique protein  IDs in both
```

* How many genes were annotated in the old and new versions?

4123

```{r}
set_both <- sans  %>%
  select(qpid) %>% 
  mutate(anno = str_split(qpid, "\\|",4)) %>% 
  mutate(Protein_ID = map(anno,nth,3)) %>% 
  mutate(Protein_ID=as.numeric(Protein_ID)) %>% 
  select(Protein_ID) %>% 
  filter(Protein_ID %in% both) %>% 
  select(Protein_ID) %>%
  distinct() %>% 
  arrange(Protein_ID) %>% pull()
```

* How many not annotated in the old are now annotated in the new?

1632

```{r}
set2 <- sans  %>%
  select(qpid) %>% 
  mutate(anno = str_split(qpid, "\\|",4)) %>% 
  mutate(Protein_ID = map(anno,nth,3)) %>% 
  mutate(Protein_ID=as.numeric(Protein_ID)) %>% 
  select(Protein_ID) %>%
  filter(Protein_ID %in% second) %>% 
  select(Protein_ID) %>%
  distinct() %>% 
  arrange(Protein_ID) %>% pull()
```

* How many not annotated in the new were annotated in the old?

2812


## How many GO terms out of the annotated in the first set not present in the second.

```{r}
set1 <- jgi %>%  
  rename(Protein_ID = `#proteinId`) %>% 
  distinct() %>%
  filter(Protein_ID %in% first) %>% 
  select(Protein_ID) %>% 
  distinct() %>% 
  pull()
```

```{r}
jgi %>%  
  rename(Protein_ID = `#proteinId`) %>% 
  select(Protein_ID,goAcc) %>% 
  filter(Protein_ID %in% first) 
```


```{r}
sessionInfo()
```

